Combining formal and distributional models of predicate semantics

### 2016-03-15 papers/acl2016/paper.tex
introduction edits
Former-commit-id: af0c43a215d6663811c735112db67026c34d5be9

How should a computer program represent the meaning of a word or a phrase?  The
answer to this question has a long and rich history.  Motivated by Harris's
distributional hypothesis~\cite{harris-1954-distributional-hypothesis}, much
work has been put into representing words as collections of the contexts in
which they appear.  Recent work with neural networks and word embeddings have
shown that these representations can be very useful for a number of natural
language processing tasks (see, e.g., the work Mikolov et
al.~\shortcite{mikolov-2013-word2vec} and papers that cite them).  This
approach to modeling semantics is appealing because word embeddings can be
trained on unlabeled text, allowing for very broad coverage, although shallow,
semantics.

@@

At the same time, other researchers have found Montague's ideas about English
as a formal language to be useful~\cite{montague-1974-english-formal-language}.
Modern semantic parsers make use of a model-theoretic semantics that represents
words and phrases as formal predicates.  These predicates are then typically
mapped onto a query over a structured knowledge base, such as
Freebase~\cite{freebase-2008-bollacker}.  This kind of modeling allows semantic
parsers to handle the compositional structure of natural language and to
leverage large, curated knowledge sources to perform various natural language
processing
tasks~\cite{zettlemoyer-2005-ccg,kwiatkowski-2013-ontology-matching,berant-2013-semantic-parsing-qa}.
However, semantic parsers are fundamentally limited by the schema of the
ontology they are trained on---a semantic parser trained with Freebase cannot
even represent words whose meaning has no correlate in the Freebase ontology,
such as ``front-runner'' and ``honcho''.

In this work we introduce models of predicate semantics that combine the
benefits of both distributional and formal approaches.  While there is no
Freebase predicate that corresponds exactly to the meaning of ``honcho'',
Freebase \emph{does} contain many predicates that can \emph{partially}
represent the word.  Our models extract features from Freebase that are
correlated with words and phrases seen in natural language text, and learns
weights for those features as additional components of a distributional model
of the word's meaning.

To demonstrate these models, we focus on the task of fill-in-the-blank natural
language queries, such as ``Italian architect \blank{}''.  Similar to many
current semantic parsers, we first convert this natural language query into a
logical form: $\lambda x. \predicate{architect}(x) \land
\predicate{architect\_N/N}(\entity{Italy},
x)$.\footnote{\predicate{architect\_N/N} represents a noun-compound
relationship between two entities mediated by the word ``architect'', as in
``Italian architect Andrea Palladio''.  This is in contrast to, e.g.,
\predicate{'s\_architect} or \predicate{architect\_of}, both of which also are
mediated by ``architect'', but through different syntax.  N/N is how
CCG~\cite{steedman-2006-ccg} represents a noun compound.} A typical semantic
parser (e.g., that of Kwiatkowski et
al.~\shortcite{kwiatkowski-2013-ontology-matching}) would then try to map this
logical form to a single Freebase query, likely involving
(\predicate{/people/person/nationality}: \entity{Italy}) and
(\predicate{/type/object/type}: \entity{/architecture/architect}).  Instead of
restricting ourselves to modeling only what can be represented as a Freebase
query, we follow Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} in learning a
model for each predicate that can give a \emph{score} to potential arguments.
The model for \predicate{architect}, e.g., should give high scores to
\entity{Andrea Palladio} and \entity{Frank Lloyd Wright}, and low scores to
\entity{Italy}, \entity{BarackObama}, etc.  However, while the models of
Krishnamurthy and Mitchell are purely distributional, using Freebase only as a
means of finding entities in text, we inject additional information from
Freebase into the models for each predicate.  The model we learn for
\predicate{architect} has a \emph{feature} capturing the fact that an entity
has the type \entity{/architecture/architect} in Freebase, and the model for
\predicate{architect\_N/N} has a \emph{feature} capturing whether the first
argument is the country of nationality of the second argument.  An overview of
the components of our model is given in \figref{overview}.

This is a new kind of predicate semantics.  It is distributional---part of the
model is still trained in a manner very similar to word2vec---yet it is also
formal.  However, instead of committing to a single Freebase query to represent
a phrase, as traditional formal approaches do, our models are flexible,
encoding many possible Freebase queries as features that can give hints to the
model.  Thus words that would be impossible to represent as a single Freebase
query can still benefit from the information contained in Freebase---our model
for ``front-runner'' learns that acceptable entities are politicians, even
though there is no exact encoding of ``front-runner'' in Freebase.

The rest of this paper is organized as follows.  In \secref{background}, we
describe the relevant background necessary to understand the models we propose,
and we situate our model in the context of related work.  In \secref{method},
we give a formal description of our models of predicate semantics.  In
\secref{evaluation}, we present experimental results on fill-in-the-blank
natural language query task.  Using web text linked to Freebase as our training
and test data, we show a 62\% relative improvement over prior work, giving a
new state-of-the-art result on this dataset.  In \secref{discussion}, we give a
detailed discussion of the benefits and drawbacks of the models we have
proposed.  We then conclude in \secref{conclusion}.

@@

This method assigns a truth probability to a predicate instance such
as \predicate{architect}(\entity{AndreaPalladio}) using probabilistic
matrix factorization. This factorization has two sets of parameters:
each category or relation has a learned $k$-dimensional embedding
$\theta$, and each entity or entity pair has a learned embedding
$\phi$. The probability assigned to a category instance $c(e)$ or
relation instance $r(e_1, e_2)$ is given by:

@@

The probability of a predicate instance is the (sigmoided) inner
product of the corresponding predicate and entity embeddings. Thus,
predicates with nearby embeddings will have similar distributions over
the entities in their denotation.

The model is trained using a query ranking objective that learns to
rank entities observed in the denotation of a logical form above
unobserved entities. The training data consists of logical form /
entity pairs $\{(\ell^i, e_^i)\}_{i=1}^n$ that are automatically
derived from an entity-linked web corpus. Training optimizes the
ranking objective function using AdaGrad, and produces the parameters
$\theta$ and $\phi$ as output. Note that this approach is similar to
modern techniques for constructing word embeddings in that it factors
a cooccurrence matrix between predicates and
entities~\cite{levy-2014-w2v-as-mf}. Given the trained predicate and
entity parameters, the system is capable of efficiently computing the
marginal probability that an entity is an element of a logical form's
denotation using approximate inference algorithms for probabilistic
databases.

>>>>

How should a computer program represent the meaning of a word or a
phrase? Recent work has suggested two vastly different approaches --
distributional vector space models and semantic parsing -- each with
its own advantages and disadvantages. On one hand, distributional
models produce a vector representation of the meaning of each word
from a large unlabeled text corpus. These models have broad coverage,
but a shallow semantics centered around word similarity. On the other
hand, semantic parsing translates natural language into logical
queries against a formal knowledge base, such as Freebase. This
approach produces a rich compositional semantics and furthermore can
leverage structured information from the knowledge base, e.g., to
answer natural language questions. However, semantic parsers have
limited coverage as they can only represent the meanings of words that
can be mapped to knowledge base predicates. A critical question is
\emph{how do we obtain the benefits of both approaches?}

% The answer to this question has a long and rich history.  Motivated by
% , much work has been put into representing words as collections of the
% contexts in which they appear.  Recent work with neural networks and
% word embeddings have shown that these representations can be very
% useful for a number of natural language processing tasks (see, e.g.,
% the work Mikolov et al.~\shortcite{mikolov-2013-word2vec} and papers
% that cite them).  This approach to modeling semantics is appealing
% because word embeddings can be trained on unlabeled text, allowing for
% very broad coverage, although shallow, semantics.

@@

% At the same time, other researchers have found Montague's ideas about English
% as a formal language to be useful~\cite{montague-1974-english-formal-language}.
% Modern semantic parsers make use of a model-theoretic semantics that represents
% words and phrases as formal predicates.  These predicates are then typically
% mapped onto a query over a structured knowledge base, such as
% Freebase~\cite{freebase-2008-bollacker}.  This kind of modeling allows semantic
% parsers to handle the compositional structure of natural language and to
% leverage large, curated knowledge sources to perform various natural language
% processing
% tasks~\cite{zettlemoyer-2005-ccg,kwiatkowski-2013-ontology-matching,berant-2013-semantic-parsing-qa}.
% However, semantic parsers are fundamentally limited by the schema of the
% ontology they are trained on---a semantic parser trained with Freebase cannot
% even represent words whose meaning has no correlate in the Freebase ontology,
% such as ``front-runner'' and ``honcho''.

Krishnamurthy and
Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary}
provides a partial answer to this question with their open-vocabulary
semantic parser. This system answers compositional, fill-in-the-blank
natural language queries against Freebase, such as ``Hollywood honcho
\blank{},'' using \emph{distributional} models of predicate
semantics. The method semantically parses text to a surface logical
form containing predicates derived from the text itself, such as
$\lambda x.\predicate{honcho}(x) \land
\predicate{honcho\_N/N}(\entity{Hollywood}, x)$, then learns the
denotations of these predicates using distributional information from
a large, entity-linked corpus. This method combines the broad coverage
of distributional approaches with the compositionality of semantic
parsing; however, it does not use structured information from Freebase
to assist with its predictions.

In this work, we introduce new models of predicate semantics that
leverage both structured information from Freebase and distributional
information. Our key observation is that, while there is no Freebase
predicate that corresponds exactly to the meaning of ``honcho,''
Freebase \emph{does} contain predicates that \emph{partially}
represent its meaning. For example, an entity that can be called
``honcho'' is likely to have type \predicate{person}. We therefore
incorporate \emph{features} based on Freebase paths into the predicate
models of K\&M and learn per-predicate weights for these features
using distributional information (\figref{overview}). This approach
effectively learns the meaning of a word as a distributional vector
plus a \emph{weighted combination of Freebase queries}, a considerably
more expressive representation than either existing semantic parsers
or K\&M.

We demonstrate our approach on the task of answering fill-in-the-blank
natural language queries, finding that our approach provides a 62\%
relative improvement in mean average precision over K\&M. (TODO: other
empirical findings?)

% Thus, we extend K\&M to incorporate such features from Freebase into
% its predicate models. Our model automatically generates such features
% and learns per-predicate weights for them using distributional
% statistics.

% In this work we introduce models of predicate semantics that combine the
% benefits of both distributional and formal approaches.  While there is no
% Freebase predicate that corresponds exactly to the meaning of ``honcho'',
% Freebase \emph{does} contain many predicates that can \emph{partially}
% represent the word.  Our models extract features from Freebase that are
% correlated with words and phrases seen in natural language text, and learns
% weights for those features as additional components of a distributional model
% of the word's meaning.

% To demonstrate these models, we focus on the task of fill-in-the-blank natural
% language queries, such as ``Italian architect \blank{}''.  Similar to many
% current semantic parsers, we first convert this natural language query into a
% logical form: $\lambda x. \predicate{architect}(x) \land
% \predicate{architect\_N/N}(\entity{Italy},
% x)$.\footnote{\predicate{architect\_N/N} represents a noun-compound
% relationship between two entities mediated by the word ``architect'', as in
% ``Italian architect Andrea Palladio''.  This is in contrast to, e.g.,
% \predicate{'s\_architect} or \predicate{architect\_of}, both of which also are
% mediated by ``architect'', but through different syntax.  N/N is how
% CCG~\cite{steedman-2006-ccg} represents a noun compound.} A typical semantic
% parser (e.g., that of Kwiatkowski et
% al.~\shortcite{kwiatkowski-2013-ontology-matching}) would then try to map this
% logical form to a single Freebase query, likely involving
% (\predicate{/people/person/nationality}: \entity{Italy}) and
% (\predicate{/type/object/type}: \entity{/architecture/architect}).  Instead of
% restricting ourselves to modeling only what can be represented as a Freebase
% query, we follow Krishnamurthy and
% Mitchell~\shortcite{krishnamurthy-2015-semparse-open-vocabulary} in learning a
% model for each predicate that can give a \emph{score} to potential arguments.
% The model for \predicate{architect}, e.g., should give high scores to
% \entity{Andrea Palladio} and \entity{Frank Lloyd Wright}, and low scores to
% \entity{Italy}, \entity{BarackObama}, etc.  However, while the models of
% Krishnamurthy and Mitchell are purely distributional, using Freebase only as a
% means of finding entities in text, we inject additional information from
% Freebase into the models for each predicate.  The model we learn for
% \predicate{architect} has a \emph{feature} capturing the fact that an entity
% has the type \entity{/architecture/architect} in Freebase, and the model for
% \predicate{architect\_N/N} has a \emph{feature} capturing whether the first
% argument is the country of nationality of the second argument.  An overview of
% the components of our model is given in \figref{overview}.
% 
% This is a new kind of predicate semantics.  It is distributional---part of the
% model is still trained in a manner very similar to word2vec---yet it is also
% formal.  However, instead of committing to a single Freebase query to represent
% a phrase, as traditional formal approaches do, our models are flexible,
% encoding many possible Freebase queries as features that can give hints to the
% model.  Thus words that would be impossible to represent as a single Freebase
% query can still benefit from the information contained in Freebase---our model
% for ``front-runner'' learns that acceptable entities are politicians, even
% though there is no exact encoding of ``front-runner'' in Freebase.
% 
% The rest of this paper is organized as follows.  In \secref{background}, we
% describe the relevant background necessary to understand the models we propose,
% and we situate our model in the context of related work.  In \secref{method},
% we give a formal description of our models of predicate semantics.  In
% \secref{evaluation}, we present experimental results on fill-in-the-blank
% natural language query task.  Using web text linked to Freebase as our training
% and test data, we show a 62\% relative improvement over prior work, giving a
% new state-of-the-art result on this dataset.  In \secref{discussion}, we give a
% detailed discussion of the benefits and drawbacks of the models we have
% proposed.  We then conclude in \secref{conclusion}.

@@

The truth probability of a predicate instance such as
\predicate{architect}(\entity{AndreaPalladio}) is learned using
probabilistic matrix factorization. This factorization has two sets of
parameters: each category or relation has a learned $k$-dimensional
embedding $\theta$, and each entity or entity pair has a learned
embedding $\phi$. The probability assigned to a category instance
$c(e)$ or relation instance $r(e_1, e_2)$ is given by:

@@

The probability of a predicate instance is the (sigmoided) inner
product of the corresponding predicate and entity embeddings. Thus,
predicates with nearby embeddings will have similar distributions over
the entities in their denotation. The parameters $\theta$ and $\phi$
are learned using a query ranking objective that optimizes them to
rank entities observed in the denotation of a logical form above
unobserved entities. Note that this approach is similar to modern
techniques for constructing word embeddings in that it factors a
cooccurrence matrix between predicates and
entities~\cite{levy-2014-w2v-as-mf}. Given the trained predicate and
entity parameters, the system is capable of efficiently computing the
marginal probability that an entity is an element of a logical form's
denotation using approximate inference algorithms for probabilistic
databases.
